# Breure et al. (2022)
# Comparing the effect of different sample conditions ... 
# Soil and Tillage research
# https://doi.org/10.1016/j.still.2021.105196

# Local library script VNIR/MIR/VNIRMIR datasets

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# LIBRARIES -----------------------------------------------------------
library(tidyverse)
library(broom)
library(modelr)
library(cowplot)
library(GeomComb)
library(clhs)
library(resemble)
library(epiR)
library(prospectr)
library(pls)
library(tidyverse)

# closeAllConnections()

```  

Read in the data
```{r, echo = FALSE, warning = FALSE}

# Reading in the data
gc()
rm(list = ls())
setwd("C:/Users/breur005/OneDrive - Wageningen University & Research/Rothamsted/Manuscript_2/Data")
list.files()

# Set seed
set.seed(123)

# Set ggplot theme
theme_set(theme_bw())

# WET CHEMISTRY -----------------------------------------------------------

WetCh_nm <- list("Wet_Chem_P", "Wet_Chem_ST", "Wet_Chem_T2", "Wet_Chem_K6")

WetCh_l <- lapply(1:length(WetCh_nm),
                  function (z)
                    read.table(paste(WetCh_nm[[z]],
                                     ".txt",
                                     sep = ""),
                               header = T,
                               sep = '\t'))

SubCh_l <- vector("list",length  = length(WetCh_l))

for(i in 1:length(WetCh_l)){
  SubCh_l[[i]] <- WetCh_l[[i]] %>% 
    dplyr::select(.,
                  ID = ID,
                  "CALCIUM" = M_Ca,
                  "MAGNESIUM" = M_Mg,
                  "SODIUM" = M_Na,
                  "SULPHUR" = M_S,
                  "CARBON" = M_C,
                  "PH" = pH,
                  "P_OLSEN" = PO4_ppm,
                  "K_NITRATE" = M_K,
                  "CLAY" = Clay_perc9,
                  "SAND" = Sand_perc,
                  "SILT" = Silt_perc9)
}



for (i in 1:length(SubCh_l)){
  print(paste("Field", i))
  print(psych::describe(SubCh_l[[i]][,2:ncol(SubCh_l[[i]])]))
}

# ------------------------------ Spectral data --------------------------------------
# -----------------------------------------------------------------------------------

# Read data
# NIR / MIR for each sample condition (field, unprocessed, air-dried, ground)
Spc <- list(list(
            list("Cal_FIELD.txt", "Cal_FIELD_ST.txt",
                 "Cal_FIELD_T2.txt", "Cal_FIELD_K6.txt"),
            list("Cal_UNP.txt", "Cal_UNP_ST.txt",
                 "Cal_UNP_T2.txt", "Cal_UNP_K6.txt"),
            list("Cal_AIR.txt", "Cal_AIR_ST.txt",
                 "Cal_AIR_T2.txt", "Cal_AIR_K6.txt"),
            list("GR_NIR.txt", "GR_NIR_ST.txt",
                 "GR_NIR_T2.txt","GR_NIR_K6.txt")
            ),
            list(
            list("Cal_FIELD_MIR.txt", "Cal_FIELD_MIR_ST.txt",
                 "Cal_FIELD_MIR_T2.txt", "Cal_FIELD_MIR_K6.txt"),
            list("Cal_UNP_MIR.txt", "Cal_UNP_MIR_ST.txt",
                 "Cal_UNP_MIR_T2.txt", "Cal_UNP_MIR_K6.txt"),
            list("Cal_AIR_MIR.txt", "Cal_AIR_MIR_ST.txt",
                 "Cal_AIR_MIR_T2.txt", "Cal_AIR_MIR_K6.txt"),
            list("GR_MIR.txt", "GR_MIR_ST.txt",
                 "GR_MIR_T2.txt","GR_MIR_K6.txt")
            ))

# Read tables
Spc_df <- lapply(1:length(Spc), 
                 function (z) 
                   lapply(1:length(Spc[[z]]), 
                                     function (s)
                                       lapply(1:length(Spc[[z]][[s]]),
                                              function (x)
                                       read.table(Spc[[z]][[s]][[x]],
                                                             header = T,
                                                             check.names = F,
                                                             sep = '\t'))))

# Write spectra as sub-dataframe (if all wavebands are retained!)
# NOTE: Only works if not any other columns have a numeric value as header
spc_subdf <- function(df){
  spc <- as.matrix(df[,8:ncol(df)])
  df <- df[,-c(8:ncol(df))]
  df$spc <- spc
  return(df)
}

Spcdf <- lapply(1:length(Spc_df), 
                function (z) 
                  lapply(1:length(Spc_df[[z]]),
                         function (s)
                           lapply(1:length(Spc_df[[z]][[s]]),
                                  function (x)
                           spc_subdf(Spc_df[[z]][[s]][[x]])
                  )
                  )
        )


# First convert wavenumber to wavelength for Tensor II NIR spectra
# I.e. given the different ranges of the spectrometers, wavenumber reduces emphasis on VIS for FieldSpec4 spectra
# I.e. decided not to convert absorbance to reflectance as there is negative absorbance values in the NSI spectra

for (j in 1:4){         # 4 fields
  colnames(Spcdf[[1]][[4]][[j]]$spc) <-   round(10000000/as.numeric(colnames(Spcdf[[1]][[4]][[j]]$spc)),1) 
  }

# Convert reflectance to absorbance to align with Tensor II spectra
for (j in 1:3){
  for (i in 1:length(Spcdf[[1]][[j]])){         # 4 fields
  Spcdf[[1]][[j]][[i]]$spc <- log10(1/Spcdf[[1]][[j]][[i]]$spc) 
  }
}

FTIR_spc_proc <- function(df){
  df$sg <- prospectr::savitzkyGolay(df$spc, p = 3, w = 11, m = 0)
  df$snv <- prospectr::standardNormalVariate(X = df$sg)
  df$d1 <- t(diff(t(df$sg), differences = 1))
  df$d2 <- t(diff(t(df$sg), differences = 2))
  df$gapD1 <- gapDer(X = df$snv, m = 1, w = 11, s = 8)
  return(df)
}

NIR_spc_proc <- function(df){
  df$sgNIR <- savitzkyGolay(df$spc, p = 3, w = 11, m = 0)
  df$snvNIR <- standardNormalVariate(X = df$sgNIR)
  df$d1 <- t(diff(t(df$sg), differences = 1))
  df$d2 <- t(diff(t(df$sg), differences = 2))
  df$gapD1NIR <- gapDer(X = df$snvNIR, m = 1, w = 11, s=31)
  #df$snv_d2 <- gapDer(X = df$snv, m = 2, w = 11)
  return(df)
}

prc_list <- list(NIR_spc_proc, FTIR_spc_proc)

Spcdf <- lapply(1:length(prc_list),
                function (s)
                  lapply(1:length(Spcdf[[s]]),
                         function (z)
                           lapply(1:length(Spcdf[[s]][[z]]),
                                  function (x)
                           prc_list[[s]](Spcdf[[s]][[z]][[x]]))))

# Remove bands functions
NIRremovebands <- function(df,uplim,lolim){
  df$spc <- df$spc[,c((colnames(df$spc) > uplim | colnames(df$spc) < lolim))]
  df$sgNIR <- df$sgNIR[,c((colnames(df$sgNIR) > uplim | colnames(df$sgNIR) < lolim))]
  df$snvNIR <- df$snvNIR[,c((colnames(df$snvNIR) > uplim | colnames(df$snvNIR) < lolim))]
  df$d1 <- df$d1[,c((colnames(df$d1) > uplim | colnames(df$d1) < lolim))]
  df$d2 <- df$d2[,c((colnames(df$d2) > uplim | colnames(df$d2) < lolim))]
  df$gapD1NIR <- df$gapD1NIR[,c((colnames(df$gapD1NIR) > uplim | colnames(df$gapD1NIR) < lolim))]
  return(df)
}

FTIRremovebands <- function(df,uplim,lolim){
  df$spc <- df$spc[,c((colnames(df$spc) > uplim | colnames(df$spc) < lolim))]
  df$sg <- df$sg[,c((colnames(df$sg) > uplim | colnames(df$sg) < lolim))]
  df$snv <- df$snv[,c((colnames(df$snv) > uplim | colnames(df$snv) < lolim))]
  df$d1 <- df$d1[,c((colnames(df$d1) > uplim | colnames(df$d1) < lolim))]
  df$d2 <- df$d2[,c((colnames(df$d2) > uplim | colnames(df$d2) < lolim))]
  df$gapD1 <- df$gapD1[,c((colnames(df$gapD1) > uplim | colnames(df$gapD1) < lolim))]
  return(df)
}



bands_l <- 1265 # 7900
bands_h <- 1460 # 6849
# Remove bands for (V)NIR
for(i in 1:4){
  for (j in 1:length(Spcdf[[1]][[i]])){
  Spcdf[[1]][[i]][[j]] <- NIRremovebands(Spcdf[[1]][[i]][[j]],
                                     bands_h,
                                     bands_l)
  }
}

bands_l <- 1789 # 5587
bands_h <- 1960 # 5102
# Remove bands for (V)NIR
for(i in 1:4){
  for (j in 1:length(Spcdf[[1]][[i]])){
  Spcdf[[1]][[i]][[j]] <- NIRremovebands(Spcdf[[1]][[i]][[j]],
                                     bands_h,
                                     bands_l)
  }
}

bands_h <- 2430
bands_l <- 2260
# Remove bands for FTIR
for(i in 1:length(Spcdf[[2]])){
  for (j in 1:length(Spcdf[[2]][[i]])){
  Spcdf[[2]][[i]][[j]] <- FTIRremovebands(Spcdf[[2]][[i]][[j]],
                                     bands_h,
                                     bands_l)
  }
}



# Rename raw NIR spc
for (i in 1:length(Spcdf[[1]])){
  for (j in 1:length(Spcdf[[1]][[i]]))
  Spcdf[[1]][[i]][[j]]$spcNIR <- Spcdf[[1]][[i]][[j]]$spc
}


# set std color
stdcol = rgb(red = .8, green = .1, blue = .1, alpha = .15)

quick_matplt <- function(df,xlab,ylab,main,col){
  tempplt <- matplot(as.numeric(colnames(df)),
                     t(df),
                     type = "l",
                     lty = 1,
                     xlab = xlab,
                     ylab = ylab,
                     col = col,
                     xlim = rev(range(as.numeric(colnames(df)))),
                     main = main)
  grid()
  return(tempplt)
}

ylab <- c("Reflectance", "Absorbance")
xlab <- c("Wavelength / nm", "Wavenumber / cm")

# Plot all fields for each processing type (i.e. field, soil moisture, air-dried)
par(mfrow = c(2,2))
Spcplt <- lapply(1:length(Spcdf), 
                 function (z)
                   lapply(1:length(Spcdf[[z]]),
                          function (s)
                            lapply(1:length(Spcdf[[z]][[s]]),
                                   function (x)
                            quick_matplt(Spcdf[[z]][[s]][[x]]$sg,
                                         xlab[z],
                                         ylab[z],
                                         "-",
                                         stdcol))))



# Merge with wet chem 
Cal_Spcdf <- lapply(1:length(Spcdf), 
                 function (z)
                   lapply(1:length(Spcdf[[z]]),
                          function (s)
                            lapply(1:length(Spcdf[[z]][[s]]),
                                   function (x)
                                     merge(SubCh_l[[x]],
                                           Spcdf[[z]][[s]][[x]],
                                           by = c("ID")))))

# Check dimensions of (V)NIR datasets
for (i in 1:length(Cal_Spcdf[[1]])){
  for (j in 1:length(Cal_Spcdf[[1]][[i]])){
    print(dim(Cal_Spcdf[[1]][[i]][[j]]))
  }
}

# Check dimensions for FTIR datasets
for (i in 1:length(Cal_Spcdf[[2]])){
  for (j in 1:length(Cal_Spcdf[[2]][[i]])){
    print(dim(Cal_Spcdf[[2]][[i]][[j]]))
  }
}


# Set Field column for T2 and K6
Cal_Spcdf[[1]][[4]][[3]]$Field <- "F3"
Cal_Spcdf[[1]][[4]][[4]]$Field <- "F4"
Cal_Spcdf[[2]][[4]][[3]]$Field <- "F3"
Cal_Spcdf[[2]][[4]][[4]]$Field <- "F4"


Cal_Compl_df <- lapply(1:length(Cal_Spcdf),
                       function (z)
                         lapply(1:length(Cal_Spcdf[[z]]),
                                function (x)
                                  do.call(rbind, Cal_Spcdf[[z]][[x]])))

test <- merge(Cal_Compl_df[[1]][[1]],
              Cal_Compl_df[[2]][[1]][,c("ID","Field","gapD1")],
              by = c("ID", "Field"))

Cal_df <- lapply(1:length(Cal_Compl_df),
                 function (z)
                   lapply(1:length(Cal_Compl_df[[z]]),
                          function (s)
                            Cal_Compl_df[[z]][[s]][interaction(Cal_Compl_df[[z]][[s]][,c("ID","Field")]) %in%  interaction(test[,c("ID","Field")]),]))

for (i in 1:length(Cal_df[[1]])){
  print(dim(Cal_df[[1]][[i]]))
}

# Run again because of incomplete Unprocessed dataset for Field 3
Cal_df <- lapply(1:length(Cal_df),
                 function (z)
                   lapply(1:length(Cal_df[[z]]),
                          function (s)
                            Cal_df[[z]][[s]][interaction(Cal_df[[z]][[s]][,c("ID","Field")]) %in%  interaction(Cal_df[[1]][[2]][,c("ID","Field")]),]))

# Check
for (i in 1:length(Cal_df[[2]])){
  print(dim(Cal_df[[2]][[i]]))
}


```

# Calibration / validation subsetting
# Spectra based method

```{r, echo = F, warning = F}


select_df <- merge(Cal_df[[2]][[4]],
                   Cal_df[[1]][[4]][,c("ID","Field","gapD1NIR")],
                   by = c("ID", "Field"))

select_df$d1_merg <- cbind(select_df$gapD1NIR,select_df$gapD1)

ken_euclid <- kenStone(X = select_df$d1_merg, 
                      k = round(nrow(select_df)*0.75), 
                      metric = "euclid")

Compl_Cal_df <- vector("list", length(Cal_df[[1]]))
for (i in 1:length(Cal_df[[1]])){
  Compl_Cal_df[[i]] <- merge(Cal_df[[2]][[i]],
                   Cal_df[[1]][[i]][,c("ID","Field","gapD1NIR")],
                   by = c("ID", "Field"))
  }

for (i in 1:length(Compl_Cal_df)){
  print(dim(Compl_Cal_df[[i]]))
}

train_df <- vector("list", length(Compl_Cal_df))
for (i in 1:length(Compl_Cal_df)){
    train_df[[i]] <- Compl_Cal_df[[i]][ken_euclid$model,]
  }


test_df <- vector("list", length(Compl_Cal_df))
for (i in 1:length(Compl_Cal_df)){
    test_df[[i]] <- Compl_Cal_df[[i]][ken_euclid$test,]
  }


# Check dimensions
for (i in 1:length(train_df)){
  print(dim(train_df[[i]]))
}

for (i in 1:length(test_df)){
  print(dim(test_df[[i]]))
}


# Write sample ID calibration set
write.table(train_df[[1]][,c("ID","Field")],
            "train_ID.txt",
            sep = '\t',
            col.names = T, row.names = F)

# Write sample ID test set
write.table(test_df[[1]][,c("ID","Field")],
            "test_ID.txt",
            sep = '\t',
            col.names = T, row.names = F)



```


# PLS modelling

# NOTE : Absolute minimum to select nr of components
# Color scatter plots by Field factor

```{r, echo = F, warning = F}

# Indexing different training datasets
train_l <- train_df
out_l <- vector("list", length(train_l))

# Indexing spectral matrices and output list
spc_matrix_l <- c("gapD1NIR", "gapD1")
spc_m_out <- vector("list", length(spc_matrix_l))


# Indexing soil variables to predict
var_l <- c("CALCIUM","CARBON", "PH",
                  "K_NITRATE", "P_OLSEN", "CLAY")

for (ii in 1:length(train_l)){
  Pls_l <- lapply(spc_matrix_l,
                function (z)
                  lapply(1:length(var_l),
                         function(s)
                           plsr(as.formula(paste(var_l[[s]],"~",as.name(z))),
                                scale = F, trace = T,
                                ncomp = 15, data = train_l[[ii]],
                                validation = "LOO")
  ))
  out_l[[ii]] <- Pls_l
  }


```

# Select components and predict on training dataset

```{r, echo = F, warning = F}

# Plot RMSE as a function of nr of components and save nr of components that minimize RMSE
min.Abs <- vector("list",length(train_l))

for(ii in 1:length(out_l)){
  for(z in 1:length(out_l[[ii]])){
  temp_comp <- lapply(1:length(out_l[[ii]][[z]]),
                  function (s)
                    which.min(
                      as.data.frame(
                        RMSEP(out_l[[ii]][[z]][[s]])$val)[1,2:16]))
  temp_comp <- ifelse(temp_comp==0,1,temp_comp)
  min.Abs[[ii]][[z]] <- temp_comp
  }
  }



test_l <- test_df
Pred_l <- vector("list",length(test_l))

# Predict most parsimonious models on test set
type_df <- c("Field", "Unprocessed", "Air-dried", "Ground")
spc_df <- c("(V)NIR", "MIR")

for(ii in 1:length(test_l)){
 Pred_l[[ii]] <- lapply(1:length(out_l[[ii]]),
                        function(xx)
                      lapply(1:length(out_l[[ii]][[xx]]),
                               function (z)
                                 data.frame(Measured = test_l[[ii]][,var_l[[z]]],
                                            Predicted = 
                   c(predict(
                     out_l[[ii]][[xx]][[z]],
                     newdata = 
                       test_l[[ii]],
                     ncomp = min.Abs[[ii]][[xx]][[z]])
                     ),
                   Soil_property = c(
                                  rep(
                                    var_l[[z]],
                                    nrow(test_l[[ii]]))),
                                Type = c(rep(type_df[[ii]],
                                           nrow(test_l[[ii]]))),
                                Spc_matrix = c(rep(spc_df[[xx]],
                                                   nrow(test_l[[ii]]))),
                   Field = test_l[[ii]]$Field
                                 )))
 }


for (ii in 1:length(type_df)){
  for (z in 1:length(spc_df)){
    for (j in 1:length(var_l)){
      Pred_l[[ii]][[z]][[j]] <-  Pred_l[[ii]][[z]][[j]] %>% 
     mutate(.,
            RMSE = signif(sqrt(mean((Predicted-Measured)^2)),2),
           RMSE_sd = sd(sqrt((Predicted-Measured)^2)),
      CCC = c(rep(epiR::epi.ccc(Predicted,Measured)$rho.c$est,
                     nrow(test_l[[ii]]))),
      SD_M = sd(Measured),
      Bias = round(mean(Predicted-Measured),2),
      SEP = sqrt(sum((Measured-Predicted)^2)/n()),
      IQ = (quantile(Measured, probs = 0.75)-quantile(Measured,probs = 0.25))) %>% 
        mutate(.,
               RPIQ = signif((IQ/SEP),2),
               RPD = signif((SD_M/RMSE),2))
    }
  }
}

Pred_train_l <- vector("list", length(train_l))

# Predict most parsimonious models on training set
for(ii in 1:length(train_l)){
 Pred_train_l[[ii]] <- lapply(1:length(out_l[[ii]]),
                        function(xx)
                          lapply(1:length(out_l[[ii]][[xx]]),
                               function (z)
                   data.frame(Measured = train_l[[ii]][,var_l[[z]]],
                              Predicted = c(predict(
                     out_l[[ii]][[xx]][[z]],
                     newdata = 
                       train_l[[ii]],
                     ncomp = min.Abs[[ii]][[xx]][[z]])),
                   Residuals = as.data.frame(out_l[[ii]][[xx]][[z]]$residuals)[,min.Abs[[ii]][[xx]][[z]]]
                     )
                 ))
  }

### --------------------------------------------------- ##
## ------------------------------ ##
## LOO-CV RMSE dataframes

LOOCV_train_l <- vector("list", length(train_l))

# Ignore row.names warning()
for(ii in 1:length(train_l)){
 LOOCV_train_l[[ii]] <- lapply(1:length(out_l[[ii]]),
                        function(xx)
                          lapply(1:length(out_l[[ii]][[xx]]),
                               function (z)
                   data.frame(Measured = train_l[[ii]][,var_l[[z]]],
                              Predicted = c(out_l[[ii]][[xx]][[z]]$validation$pred),
                   Soil_property = c(var_l[[z]]),
                                Type = c(type_df[[ii]]),
                                Spc_matrix = c(spc_df[[xx]]),
                   Ncomp = min.Abs[[ii]][[xx]][[z]],
                   Field = train_l[[ii]]$Field
                     )
                 ))
  }

# Compute prediction metrics
for (ii in 1:length(type_df)){
  for (z in 1:length(spc_df)){
    for (j in 1:length(var_l)){
      LOOCV_train_l[[ii]][[z]][[j]] <-  LOOCV_train_l[[ii]][[z]][[j]] %>% 
     mutate(.,
            RMSE = signif(sqrt(mean((Predicted-Measured)^2)),2),
           RMSE_sd = sd(sqrt((Predicted-Measured)^2)),
      CCC = c(epiR::epi.ccc(Predicted,Measured)$rho.c$est),
      SD_M = sd(Measured),
      Bias = round(mean(Predicted-Measured),2),
      SEP = sqrt(sum((Measured-Predicted)^2)/n()),
      IQ = (quantile(Measured, probs = 0.75)-quantile(Measured,probs = 0.25))) %>% 
        mutate(.,
               RPIQ = signif((IQ/SEP),2),
               RPD = signif((SD_M/RMSE),2))
    }
  }
}

```

# Combining predictions

# TESTING DATASET

```{r, echo = F, warning=F}

# ---------------------------
# ---------------------------

# NIRMIR model averaging
NIRMIR_Pred_l <- vector("list", length(train_l))
for (ii in 1:length(Pred_l)){
  for (xx in 1:length(var_l)){
    NIRMIR_Pred_l[[ii]][[xx]] <- cbind(
      Pred_l[[ii]][[1]][[xx]][,c("Predicted")],
      Pred_l[[ii]][[2]][[xx]][,c("Predicted")])                                 }
}


# TRAINING DATASET

# NIRMIR model averaging
NIRMIR_Pred_train_l <- vector("list", length(train_l))
for (ii in 1:length(Pred_train_l)){
  for (xx in 1:length(var_l)){
    NIRMIR_Pred_train_l[[ii]][[xx]] <- cbind(
      Pred_train_l[[ii]][[1]][[xx]][,c("Predicted")],
      Pred_train_l[[ii]][[2]][[xx]][,c("Predicted")])                                 }
}
```

# Forecomb object indexing

```{r, echo = F, warning = F}

# NIR MIR
FORC_NIRMIR <- vector("list", length(train_l))
for (ii in 1:length(Pred_train_l)){
  for (xx in 1:length(var_l)){
    FORC_NIRMIR[[ii]][[xx]] <- 
      foreccomb(train_l[[ii]][,var_l[[xx]]],
                NIRMIR_Pred_train_l[[ii]][[xx]],
                #test_l[[ii]][,var_l[[xx]]],
                newpreds = NIRMIR_Pred_l[[ii]][[xx]])
  }
}


```

# OLS Regression

```{r, echo = F, warning = F}

OLS_Results <- lapply(1:length(FORC_NIRMIR),
                          function (z)
                            lapply(1:length(var_l),
                                   function (j)
                                     comb_OLS(
                                  FORC_NIRMIR[[z]][[j]])))


```

# Create dataframes for OLS result (forecomb_res objects)

```{r, echo = F, warning = F}

OLS_l_df <- lapply(1:length(OLS_Results),
                        function(xx)
                          lapply(1:length(OLS_Results[[xx]]),
                               function (z)
                                 data.frame(Measured = 
                              test_l[[xx]][,var_l[[z]]],
                                            Predicted = 
                                OLS_Results[[xx]][[z]]$Forecasts_Test,
                   Soil_property = c(
                                  rep(
                                    var_l[[z]],
                                    nrow(test_l[[xx]]))),
                                Type = c(rep(type_df[[xx]],
                                           nrow(test_l[[xx]]))),
                                Spc_matrix = c(rep("(V)NIRMIR",
                                                   nrow(test_l[[xx]]))),
                   Field = test_l[[xx]]$Field
                                 )))



for (ii in 1:length(OLS_l_df)){
  for (z in 1:length(OLS_l_df[[ii]])){
      OLS_l_df[[ii]][[z]] <-  OLS_l_df[[ii]][[z]] %>% 
    mutate(.,
            RMSE = signif(sqrt(mean((Predicted-Measured)^2)),2),
           RMSE_sd = sd(sqrt((Predicted-Measured)^2)),
      CCC = c(rep(epiR::epi.ccc(Predicted,Measured)$rho.c$est,
                     nrow(test_l[[ii]]))),
      SD_M = sd(Measured),
      Bias = round(mean(Predicted-Measured),2),
      SEP = sqrt(sum((Measured-Predicted)^2)/n()),
      IQ = (quantile(Measured, probs = 0.75)-quantile(Measured,probs = 0.25))) %>% 
        mutate(.,
               RPIQ = signif((IQ/SEP),2),
               RPD = signif((SD_M/RMSE),2))

    }
  }


```

# Merge NIR, MIR, NIRMIR dataframes in long format

```{r, echo = F, warning = F}

# Bind at any depth function
depth <- function(this) ifelse(is.list(this), 1L + max(sapply(this, depth)), 0L)

bind_at_any_depth <- function(l) {
  if (depth(l) == 2) {
    return(bind_rows(l))
  } else {
    l <- map_depth(l, depth(l) - 2, bind_rows)
    bind_at_any_depth(l)
  }
}

Pred_l_df_unnested = bind_at_any_depth(Pred_l)
OLS_l_df_unnested = bind_at_any_depth(OLS_l_df)

Compl_pred <- rbind(Pred_l_df_unnested,OLS_l_df_unnested)

# Check frequency
table(Compl_pred$Field)

```

# Create Figures 1--5

```{r, echo = F, warning = F}

# Plot predicted versus measured for each soil property
cols <- c("Soil_property", "Type", "Spc_matrix", "Field")

Compl_pred[cols] <- lapply(Compl_pred[cols], factor)

glimpse(Compl_pred)

# Only include NIR, MIR and NIRMIR within prediction plots
Compl_pred <- Compl_pred[Compl_pred$Spc_matrix == "(V)NIR" |
                           Compl_pred$Spc_matrix == "MIR" |
                           Compl_pred$Spc_matrix == "(V)NIRMIR", ]

# Fix order for facetting
Compl_pred$Spc_matrix <- factor(Compl_pred$Spc_matrix, levels = c("(V)NIR", "MIR", "(V)NIRMIR"))

Compl_pred$Type <- recode(Compl_pred$Type,
                                 "Field" = "In-situ",
                                "Unprocessed" = "Unprocessed",
                                "Air-dried" = "Air-dried",
                          "Ground" = "Milled")

# Fix order facetting
Compl_pred$Type <- factor(Compl_pred$Type, levels = c("In-situ", "Unprocessed", "Air-dried", "Milled"))

var_labels <- c("Calcium / mg kg", "Carbon / %",
                "pH", "K / mg kg)","P / mg kg",
                "Clay / %")

RMSE_lab <- c("(mg kg) :", "(%) :", "(-) :","(mg kg) :","(mg kg) :", "(%) :")


# Subset to export  
Exp_pred_soil <- Compl_pred[Compl_pred$Soil_property != "CALCIUM",] # Exclude calcium
Exp_pred_soil <- Exp_pred_soil[Exp_pred_soil$Type == "Milled",]
Exp_pred_soil <- Exp_pred_soil[Exp_pred_soil$Spc_matrix == "(V)NIRMIR",]

write.table(Exp_pred_soil,
            "Local_pred.txt", col.names = T, row.names = F, sep = '\t')


Compl_pred_soil <- lapply(1:length(var_l),
                          function (z)
                            Compl_pred[Compl_pred$Soil_property == var_l[[z]],])

# Facet wrap approach, free scales
FacetLims_list <-  lapply(1:length(Compl_pred_soil),
                          function (z)
                            Compl_pred_soil[[z]] %>%
                                  group_by(.,Soil_property,Type,Spc_matrix) %>%
                                  summarize(min = min(Measured, Predicted),
                                            max = max(Measured, Predicted)) %>%
                                  gather(.,range, Measured, -Soil_property,
                                         -Type, -Spc_matrix) %>%
                                  mutate(Predicted = Measured, range = NULL))

# Rewrite limit for facet
FacetLims_list[[5]][24,4] <- 80
FacetLims_list[[5]][24,5] <- 80


# Plotting
GGpl_l_soil_property <- lapply(1:length(var_l),
                               function (i) 
                                 ggplot(Compl_pred_soil[[i]],
                                             aes(Measured,Predicted))+
                              geom_point()+
                              geom_abline(intercept = 0, slope = 1)+
                                xlab(paste('Measured', var_labels[[i]]))+
                                ylab(paste('Predicted', var_labels[[i]]))+
                                facet_grid(Type~Spc_matrix)+
                                geom_blank(data = FacetLims_list[[i]])+
                                geom_text(x = -Inf, y = Inf, 
                                        label = paste("RMSE ",RMSE_lab[i], 
                                                      signif(Compl_pred_soil[[i]]$RMSE, 2)), 
                                        parse = F, hjust = 0, vjust = 1.50, size = 3, color = "black")+
                                geom_text(x = -Inf, y = Inf, 
                                        label = paste("CCC :", 
                                                      signif(Compl_pred_soil[[i]]$CCC, 2)), 
                                        parse = F, hjust = 0, vjust = 3.00, size = 3, color = "black")+
                                geom_text(x = -Inf, y = Inf, 
                                        label = paste("Bias :", 
                                                      signif(Compl_pred_soil[[i]]$Bias, 2)), 
                                        parse = F, hjust = 0, vjust = 4.25, size = 3, color = "black")+
                                geom_text(x = -Inf, y = Inf, 
                                        label = paste("RPIQ :", 
                                                      signif(Compl_pred_soil[[i]]$RPIQ, 2)), 
                                        parse = F, hjust = 0, vjust = 5.75, size = 3, color = "black"))


# dir.create("Figures")
for(i in 1:length(GGpl_l_soil_property)){
  tiff(filename = paste("./Figures/Model_ave_pred_",var_l[[i]],".tif", sep = ""),
      height = 7, width = 6, units = "in", pointsize = 1/400, res = 400)
  print(GGpl_l_soil_property[[i]])
  dev.off()
  }




```

# CCC vs. RPIQ

```{r, echo = F, warning = F}

do.call(rbind, Compl_pred_soil) %>%
  filter(.,
         Soil_property != "CALCIUM") %>% 
  ggplot(aes(x = CCC, y = RPIQ, col = Soil_property))+
  geom_point()+
  geom_line()+
  xlab("CCC / (-)")+
  ylab("RPIQ / (-)")+
  labs(col="Soil property")


```

# Graphical abstract

```{r, echo = F, warning = F}

# Subset VNIRMIR data
subDF <- Compl_pred[Compl_pred$Spc_matrix == "(V)NIRMIR",]

# Split based on subsetted typed
splsubDF <- split(subDF, subDF$Type)

# Group and summarize
sumsplsubDF <- lapply(splsubDF,
                      function (x) {
                        x %>% 
                        group_by(.,Soil_property) %>% 
                        summarise(RMSE = mean(RMSE),
                                  RPIQ = mean(RPIQ),
                                  Bias = mean(Bias))
                        }
                      )

# Check prediction metrics MIR
for (i in 1:length(sumsplsubDF)){
  sumsplsubDF[[i]]$RPIQ_ratio <- sumsplsubDF[[i]]$RPIQ/sumsplsubDF[[4]]$RPIQ
  print(splsubDF[[i]]$Type[1])
  print(sumsplsubDF[[i]])
  
}

# Read prediction metrics from spectral libraries
subF4 <- read.table("RPIQRMSE_spiked.txt",
                    header = T, sep = '\t')

# Make df for plotting
subDF2 <- subDF %>% 
  select(., Soil_property, Type, RMSE, RPIQ, Bias) %>% 
  setNames(., c("Soil_property","Dataset","RMSE","RPIQ","Bias")) %>% 
  filter(Soil_property != "CALCIUM") %>% 
  mutate(.,
         Soil_property = recode(Soil_property,
                                CARBON = "Carbon (%)",
                                PH = "pH",
                                CLAY = "Clay (%)",
                                K_NITRATE = "K (mg/kg)",
                                P_OLSEN = "P (mg/kg)"),
         Dataset = recode(Dataset,
                       "In-situ" = "FD (In-situ)",
                       Unprocessed = "FD (Unprocessed)",
                       "Air-dried" = "FD (Air-dried)",
                       Milled = "FD (Milled)")) %>%
  bind_rows(.,subF4) %>% 
  group_by(Soil_property, Dataset) %>% 
  summarise(n = sum(RMSE),
            m = sum(RPIQ),
            z = abs(sum(Bias))) %>%
  mutate("RMSE (V)NIRMIR predictions" = n / sum(n),
         "RPIQ (V)NIRMIR predictions" = m / sum(m),
         "Bias (V)NIRMIR predictions" = z / sum(z)) %>% 
  gather(.,
         key = "key",
         value = "percentage",
         -Soil_property, -Dataset, -n, -m, -z)

# Rewrite factor levels
subDF2$Dataset <- factor(subDF2$Dataset,
                         levels(factor(subDF2$Dataset))[c(1,2,3,5,8,6,7,4)])

# png(filename = "./Figurs/Graphical_abstract.png", 
#     units = "in", width = 12, height = 4, res = 400)
subDF2 %>% 
  mutate(.,
         Soil_property = recode(Soil_property,
                                "Carbon (%)" = "SOC",
                                "pH" = "pH",
                                "K (mg/kg)" = "KExch",
                                "P (mg/kg)" = "POlsen",
                                "Clay (%)" = "Clay")) %>%
  #filter(Dataset != "Field-scale (Milled)") %>% 
  ggplot(., aes(x = Soil_property, y = percentage, 
                fill = Dataset, group = Dataset))+
  geom_area(alpha = 0.6, size = 1, color = "black")+
  facet_grid(~key)+
  ylab("%")+
  coord_cartesian(expand = FALSE)+
  scale_y_continuous(limits = c(0,1))+
  xlab(NULL)+
  theme(text = element_text(size = 14),
        panel.spacing.x = unit(2.5, "lines"),
        strip.text.x  = element_text(size = 14))+
  scale_fill_manual(values = c("slategray1",
                               "orange",
                               "forestgreen",
                               "lemonchiffon",
                               "lightcoral",
                               "darkorchid",
                               "burlywood4",
                               "turquoise4"))
# dev.off()

```

